Accident prediction in time and space

### 2019-04-15 presentation/pres.tex
add figures and results

\begin{itemize}
\item[] \textbf{Context}:
  \begin{itemize}
    \item Road accidents implies millions of people deaths and injuries every year.
		\item Human and economic impact on society.
    \item Master’s thesis of Antoine Hébert in a different context.
  \end{itemize}

\item[] \textbf{Our goals}:
  \begin{itemize}
    \item Dataset analysis.
  	\item Finding key insights in accident prediction.
  	\item Identifying issues related to class imbalance and geo-spatial data.
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item[] \textbf{National Collision Database (NCDB)}:
  \begin{itemize}
  	\item All vehicle collisions reported by the police from 2012 to 2017 in Montreal.
  	\item Date and time, severity, number of death and injury, number of vehicles, localization, speed limit etc.
  \end{itemize}
\item[] \textbf{National Road Network}:
  \begin{itemize}
    \item Shape of the road, how roads are
connected together, the mail addresses on the street, the number of lanes in each direction and the
type of pavement.
  \end{itemize}
\item[] \textbf{Climate data - Hourly Data Report}:
  \begin{itemize}
    \item For each position and hour, the dataset provides the temperature, the humidity, the wind speed and direction, the visibility, the atmospheric pressure etc.
  \end{itemize}
\end{itemize}

@@

\begin{frame}
	\frametitle{Generating the negative samples}
	\begin{itemize}
    \item Every time an accident did not happen is considered a negative sample.
		\item We extracted a subsample from the cartesian product between all dates (between 01/01/2012 and 31/12/2017) and all roads.
		\item Generation of 5 million negative samples for about 200 000 positive samples.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Some feature engineering}
	\begin{itemize}
		\item Computation of the distances between roads and accidents using their GPS coordinates (taking into account earth curvature)
		\item Transformation of the dates and times into a cyclic format
		\item Extrapolation of missing points on the roads (defined by a limited number of points in a KML format)
    \item Interpolation of weather statistics at a given point by averaging data collected from several stations
    \item etc...
	\end{itemize}
\end{frame}

@@

  \item[] \textbf{BRF versus WRF}:
    \begin{itemize}
      \item No clear winner ! But...
      \item WRF - More vulnerable to noise on minority class (\cite{Chen2004})
      \item BRF - More computationally efficient because use less data (subsampling) (\cite{Chen2004})
      \item BRF - Easier to implement in Spark (see next slide)
    \end{itemize}

@@

\begin{frame}
\frametitle{Experimenting with our BRF implementation}
\begin{itemize}
  \item Area Under PR : 99,3\% for WRF VS 99,7\% for BRF
  \item F1 score : 99\% for WRF VS 98\% for BRF (which proves that F1 score can be misleading in the context of class imbalance)

@@

\centering
\includegraphics[height=3.5cm, keepaspectratio]{Figures/BRF.png}
\includegraphics[height=3.5cm, keepaspectratio]{Figures/WRF.png}

@@

\begin{frame}
	\frametitle{Dataset generation}
	\begin{itemize}
		\item Using slurm-based (scheduler) cluster to increase the dataset generation's speed.
    \item Tricks to increase the generation speed and reduce the memory consumption.
	\end{itemize}
\centering
\includegraphics[height=3.5cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples.png}
\includegraphics[height=3.5cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples_after.png}
\end{frame}

\begin{frame}
	\frametitle{Spark VS Dask}
	\begin{itemize}
    \item[] \textbf{Advantages of using Dask over Spark}
      \begin{itemize}
      	\item Mimic the pandas dataframe API
        \item Lighter than Spark
      	\item Designed to be more flexible in terms of applications and algorithms
      \end{itemize}
    \item[] \textbf{Why we switched for Spark}
      \begin{itemize}
        \item We found ourselves writing a lot more code using Dask than using Spark
        \item We found Spark easier to use for processing the dataset in batch
        \item Spark's map-reduce and shuffling operations are very handy
        \item Spark ML pipelines allow to create processing pipelines easily
      \end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Comparison between the algorithms}
	\begin{itemize}
		\item Random forest
		\item Balanced random forest
		\item XGBoost
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
    \item We found some key features that can explain the occurence of accidents.
    \item We identified some issues related to class imbalance and big data analytics.
    \item We implemented BRF in Spark.
    \item Future work would include tuning the XGBoost classifier, testing lightGBM classifier, adding some datasets like a road works dataset.
    \item The code can be found on Github at https://github.com/GTimothee/accident-prediction-montreal
	\end{itemize}
\end{frame}

>>>>

\begin{itemize}
\item[] \textbf{Context}:
  \begin{itemize}
    \item Road accidents result in millions of people deaths and injuries every year.
		\item Human and economic impact on society.
    \item Master’s thesis of Antoine Hébert in a different context.
  \end{itemize}

\item[] \textbf{Our goals}:
  \begin{itemize}
    \item Dataset assembling.
    \item Predict the risk of road accident at a given time on a given road.
  	\item Finding key insights in accident prediction.
  	\item Identifying issues related to class imbalance and geo-spatial data.
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item[] \textbf{National Collision Database (NCDB)}:
  \begin{itemize}
  	\item All vehicle collisions reported by the police from 2012 to 2017 in Montreal.
  	\item Date and time, severity, number of death and injury, number of vehicles, localization, speed limit etc.
  \end{itemize}
\item[] \textbf{National Road Network}:
  \begin{itemize}
    \item Shape of the road, how roads are
connected together, the mail addresses on the street, etc.
  \end{itemize}
\item[] \textbf{Climate data - Hourly Data Report}:
  \begin{itemize}
    \item The dataset provides the temperature, the humidity, the wind speed and direction, the visibility, the atmospheric pressure etc. at different weather stations in Montreal.
  \end{itemize}
\end{itemize}

@@

\begin{frame}
	\frametitle{Generating the examples}
  \begin{itemize}
  \item[] \textbf{Positive samples}:
    \begin{itemize}
      \item Identification of road for each accident: Computation of the distances between roads and accidents using their GPS coordinates (taking into account earth curvature)
		  \item Extrapolation of missing points on the roads for more precise road identification
    \end{itemize}
  \item[] \textbf{Negative samples}:
    \begin{itemize}
      \item Every time an accident did not happen is considered a negative sample.
  		\item We extracted a subsample from the cartesian product between all dates (between 01/01/2012 and 31/12/2017) and all roads.
  		\item Generation of 5 million negative samples for about 200 000 positive samples.
    \end{itemize}

\begin{frame}
	\frametitle{Some feature engineering}
	\begin{itemize}
		\item Transformation of the dates and times into a cyclic format
    \item Interpolation of weather statistics at a given point by averaging data collected from several stations
    \item Road features: street type and length
	\end{itemize}
\end{frame}

@@

  \item[] \textbf{BRF versus WRF}:
    \begin{itemize}
      \item No clear winner in terms of prediction power ! But...
      \item WRF - More vulnerable to noise on minority class (\cite{Chen2004})
      \item BRF - More computationally efficient because use less data (subsampling) (\cite{Chen2004})
      \item BRF - Easier to implement in Spark (see next slide)
    \end{itemize}

@@

\begin{frame}
\frametitle{Experimenting with our BRF implementation}
\begin{itemize}
  \item Area Under PR : 98,1\% for RF VS 98,7\% for BRF
  \item F1 score : 99.7\% for RF VS 98.8\% for BRF (which proves that F1 score can be misleading in the context of class imbalance)
  \item With BRF, we detect more accidents but it increases the false positive rate.

@@

\centering
\includegraphics[height=3.0cm, keepaspectratio]{Figures/RF.png}
\includegraphics[height=3.0cm, keepaspectratio]{Figures/BRF.png}

@@

\begin{frame}
	\frametitle{Dataset generation}
	\begin{itemize}
		\item Using cluster to increase the dataset generation's speed.
    \item Tricks to increase the generation speed and reduce the memory consumption.
	\end{itemize}
\centering
\includegraphics[height=5.0cm, keepaspectratio]{Figures/timeVSsamples_generatepositivesamples.png}
\end{frame}

\begin{frame}
	\frametitle{Spark VS Dask}
	\begin{itemize}
    \item[] \textbf{Advantages of using Dask over Spark}
      \begin{itemize}
      	\item Mimic the pandas dataframe API
        \item Lighter than Spark
      	\item Designed to be more flexible in terms of applications and algorithms
      \end{itemize}
    \item[] \textbf{Why we switched for Spark}
      \begin{itemize}
        \item We found ourselves writing a lot more code using Dask than using Spark
        \item We found Spark easier to use for processing the dataset in batch
        \item Spark's map-reduce and shuffling operations are very handy
      \end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Results}
	\begin{itemize}
		\item Random forest: 15.3\% area under PR curve
		\item Balanced random forest: 17.2\% area under PR curve
	\end{itemize}
\centering
\includegraphics[height=4cm, keepaspectratio]{Figures/rf_results.png}
\includegraphics[height=4cm, keepaspectratio]{Figures/nohyperparam_brf.png}
\end{frame}

@@

\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
    \item We found some key features that can explain the occurence of accidents.
    \item We identified some solution to deal with class imbalance and big data-related issues.
    \item We implemented BRF in Spark.
    \item Future work would include tuning the XGBoost classifier, adding some datasets like a road works dataset.
    \item The code can be found on Github at https://github.com/GTimothee/accident-prediction-montreal
	\end{itemize}
\end{frame}
